Evaluating the Effects of Tracker Reliability and Field of View on a Target Following Task in Augmented Reality

### 2009-06-25 trunk/Paper/paper.tex

Livingston and Ai studied the effects of various sources of registration error with an AR simulation similar to our own \cite{4637329}.  Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.  A white box was continuously visible, even when the car itself was occluded by a building in the environment.  Other virtual cars and their associated white boxes acted as distractors.  At specific times during the experiment, the simulation would freeze and the participant was asked to align the center of their view (indicated by cross-hairs) on the location of the correct car and press a key.  While the effects of several types of error were studied in this work, the effect of registration dropouts was not examined.

@@

\begin{figure*}[ht!]
	\centering
	\includegraphics[width=7in]{figures/triple_view.png}
	\caption{Sample views of AR simulation.  Transparent red rectangles overlay the people, indicating their location.  \emph{Left:} A user view of Top Hat (non-occluded) in the doorway.   \emph{Middle:} Three distractor people are visible through a large window.  \emph{Right:} A typical user view of a corner occluding several people.}
\end{figure*}

@@

The parameters of our experiment are divided between the ``real'' components of immersion, such as the field of view of the HMD, and the ``augmented'' components of immersion, such as the field of view of the simulated AR display, and the performance of the tracking sensor.  We introduce periods of sensor dropouts, where the augmented overlays disappear, to simulate the effect of dropouts in a real tracking system.  For example, such dropouts might occur with a magnetic sensor near interfering material.

@@

Decreased reliability in the head-based rendering, caused by periods of sensor dropout, adversely affects performance.  However, the performance decrease levels off at one second dropouts.  Preliminary analysis also shows no significant difference between having no dropouts and half second dropouts.  This suggests that below a certain length of dropout time, good performance may be maintained, given a sufficiently large FOV.  This threshold on dropouts would change depending on the task at hand, but we expect the same pattern would arise in other target following scenarios.

>>>>

Livingston and Ai studied the effects of various sources of registration error with an AR simulation similar to our own \cite{4637329}.  Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.  A white box was continuously visible, even when the car itself was occluded by a building in the environment.  Other virtual cars and their associated white boxes acted as distractors.  At specific times during the experiment, the simulation would freeze and the participant identified the location of the correct car.

@@

\begin{figure*}[ht!]
	\centering
	\includegraphics[width=5in]{figures/triple_view.png}
	\caption{Sample views of AR simulation.  Transparent red rectangles overlay the people, indicating their location.  \emph{Left:} A user view of Top Hat (non-occluded) in the doorway.   \emph{Middle:} Three distractor people are visible through a large window.  \emph{Right:} A typical user view of a corner occluding several people.}
\end{figure*}

@@

The parameters of our experiment are divided between the ``real'' components of immersion, such as the field of view of the HMD, and the ``augmented'' components of immersion, such as the field of view of the simulated AR display, and the performance of the tracking sensor.  We introduce periods of sensor dropouts, where the augmented overlays disappear, to simulate the effect of dropouts in a real tracking system.  For example, such dropouts might occur with a magnetic sensor near interfering material.  Different display options are available in this case; for example we could use the last known sensor reading, or predict future values.
%However, these options could have a positive or negative effect on performance, depending on the circumstances.
However, we determined that hiding the AR overlay during sensor dropouts would have the least effect on performance.

@@

Decreased reliability in the head-based rendering, caused by periods of sensor dropout, adversely affects performance.  However, the performance decrease levels off at one second dropouts.  Preliminary analysis also shows no significant difference between having no dropouts and half second dropouts.  This suggests that below a certain length of dropout time, good performance may be maintained, given a sufficiently large FOV.%  This threshold on dropouts would change depending on the task at hand, but we expect the same pattern would arise in other target following scenarios.
